---
title: "Analysis and sizing: trade-offs, curves and diagnostics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis and sizing: trade-offs, curves and diagnostics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4.2)
has_ggplot <- requireNamespace("ggplot2", quietly = TRUE)
```

```{r setup}
library(rharv)
p <- precip_pi$value
d <- precip_pi$date
```

The daily simulation (`rh_simulate()`) is the engine; the functions below sweep
its parameters to build decision tools. Two reliability metrics are
available throughout and are interchangeable via the `metric` argument:
`"attendance_pct"` (volumetric: share of demand volume met) and
`"reliability_pct"` (temporal: share of days without a shortfall).

> Two-parameter grids run hundreds of simulations. For interactive exploration
> use `climatology = TRUE`, which simulates on the day-of-year mean (about 70x
> faster); confirm final numbers on the full series.

## Roof, tank or demand: comparing the levers

`rh_grid()` sweeps two levers; `rh_plot_tradeoff()` draws the resulting surface
with iso-performance contours. Here, area vs capacity at a fixed demand:

```{r tradeoff, eval = has_ggplot}
g <- rh_grid(
  p, base = list(demand = 6.6, runoff = 0.85, efficiency = 1),
  x = "area", x_values = seq(800, 8000, length.out = 25),
  y = "capacity", y_values = seq(100, 1500, length.out = 25),
  metrics = "attendance_pct", dates = d, climatology = TRUE
)
rh_plot_tradeoff(g, breaks = c(60, 70, 80, 90, 100))
```

Reading across the contours shows whether enlarging the roof (move right) or the
tank (move up) reaches a target guarantee with less change.

From a single operating point, `rh_plot_levers()` compares the gain from each
lever applied alone (here +25% area, +25% capacity, -10% demand):

```{r levers, eval = has_ggplot}
rh_plot_levers(
  p, base = list(demand = 6.6, area = 4170, capacity = 400,
                 runoff = 0.85, efficiency = 1),
  dates = d, climatology = TRUE
)
```

## Storage-Yield-Reliability and the guarantee frontier

Reliability against capacity, one line per demand (the classic SYR view):

```{r syr, eval = has_ggplot}
gs <- rh_grid(
  p, base = list(area = 4170, runoff = 0.85, efficiency = 1),
  x = "capacity", x_values = seq(100, 1500, length.out = 20),
  y = "demand", y_values = c(3, 5, 7, 9),
  metrics = "reliability_pct", dates = d, climatology = TRUE
)
rh_plot_syr(gs)
```

The zero-deficit frontier (guaranteed demand for each capacity):

```{r frontier}
rh_guarantee_curve(
  p, base = list(area = 4170, runoff = 0.85, efficiency = 1),
  vary = "capacity", values = seq(100, 1000, by = 150), target = "demand",
  dates = d, climatology = TRUE
)
```

A non-dimensional design curve (transferable across sites): demand and capacity
are divided by the mean annual inflow.

```{r design, eval = has_ggplot}
n_years <- length(unique(format(d, "%Y")))
inflow <- sum(rh_available_volume(p, area = 4170, runoff = 0.85, efficiency = 1)) / n_years
gd <- rh_grid(
  p, base = list(area = 4170, runoff = 0.85, efficiency = 1),
  x = "demand", x_values = seq(2, 12, length.out = 20),
  y = "capacity", y_values = seq(100, 1500, length.out = 20),
  metrics = "reliability_pct", dates = d, climatology = TRUE
)
rh_plot_design_curve(gd, annual_inflow = inflow)
```

## Design curves: the stage-area-volume analogue

Iso-guarantee curves relate two levers at fixed guarantee levels (here area vs
the reservoir capacity needed, at fixed demand): along each line you trade roof
for tank without changing the guarantee.

```{r iso, eval = has_ggplot}
iso <- rh_iso_curve(
  p, base = list(demand = 6.6, runoff = 0.85, efficiency = 1),
  x = "area", x_values = seq(1000, 7000, length.out = 12),
  y = "capacity", levels = c(70, 80, 90, 100),
  dates = d, climatology = TRUE
)
rh_plot_iso(iso)
```

The closest analogue to the reservoir stage-area-volume curve puts the guarantee
on the x-axis, with the required capacity and area on a dual y-axis.

```{r resource, eval = has_ggplot}
rc <- rh_resource_curve(
  p, base = list(demand = 6.6, area = 4170, capacity = 400,
                 runoff = 0.85, efficiency = 1),
  levels = seq(50, 100, by = 5), dates = d, climatology = TRUE
)
rh_plot_resource_curve(rc)
```

## Diagnostics

```{r sim}
sim <- rh_simulate(p, demand = 6.6, area = 4170, capacity = 400,
                   runoff = 0.85, efficiency = 1)
```

Reservoir behaviour over the first three years:

```{r behaviour, eval = has_ggplot}
n <- 3 * 365
sim3 <- rh_simulate(p[seq_len(n)], demand = 6.6, area = 4170, capacity = 400,
                    runoff = 0.85, efficiency = 1)
rh_plot_behaviour(sim3, dates = d[seq_len(n)])
```

Monthly water balance and the failure calendar (when shortfalls happen):

```{r calendar, eval = has_ggplot, fig.height = 5}
rh_plot_monthly_balance(sim, dates = d)
rh_plot_failure_calendar(sim, dates = d)
```

## Rainfall variability and scenarios

The spread of annual rainfall (mean with 10th-90th percentile band):

```{r spread, eval = has_ggplot}
rh_plot_spread(rh_series_spread(precip_pi, by = "year"))
```

Stress test by ordering the years dry-to-wet, and compare to the chronological
series:

```{r scenarios}
dry_first <- rh_scenarios_from_years(precip_pi, order = "increasing")
sim_stress <- rh_simulate(dry_first$value, demand = 6.6, area = 4170,
                          capacity = 400, runoff = 0.85, efficiency = 1)
rh_compare(list(chronological = sim, dry_to_wet = sim_stress))[
  , c("name", "attendance_pct", "reliability_pct", "deficit_total")]
```

## Interactive exploration

For an interactive version of all of this, launch the Shiny app:

```{r app, eval = FALSE}
rh_explore()
```
