---
title: "Worked example: rainwater harvesting at IFPB-PI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Worked example: rainwater harvesting at IFPB-PI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
has_ggplot <- requireNamespace("ggplot2", quietly = TRUE)
```

```{r setup}
library(rharv)
```

## Overview

This vignette walks through a complete rainwater-harvesting analysis with
`rharv`, applied to the Federal Institute of Paraíba, Campus Princesa Isabel
(IFPB-PI), a campus in the Brazilian semi-arid region. It shows how to go from a
daily rainfall series to system metrics and design figures using the package's
generic functions. Anything specific to this application (the campus parameters
and the scenarios) is defined in this vignette rather than shipped with the
package.

The analysis uses a daily water-balance model: the captured volume
(`Vdisp = P * A * C * eta`) feeds the reservoir mass balance, simulated day by
day. In this application the runoff coefficient and the system efficiency are
taken together as a single `C = 0.85` (`runoff = 0.85`, `efficiency = 1`).
Volumes are in cubic metres (m³) and the reservoir starts full
(`initial = capacity`). The daily-simulation approach and the available-volume
relation follow established references (Souza, 2015; ABNT NBR 15527).

## Data

`rharv` ships the cleaned daily precipitation series of the Princesa Isabel gauge
(`precip_pi`) and the campus roof areas (`areas_pi`).

```{r data}
str(precip_pi)
areas_pi
area_total <- sum(areas_pi$area_m2)   # total catchment area (m2)
area_total
```

Day-of-year precipitation climatology (the "precipitation matrix"):

```{r climatology, eval = has_ggplot}
rh_plot_climatology(precip_pi)
```

## Demand

The campus serves 1089 people at about 6.03 L/person/day:

```{r demand}
demand <- rh_daily_demand(population = 1089, per_capita_l = 6.03)  # m3/day
demand
```

## Scenarios

We compare two precipitation scenarios, built by sub-setting the series:

* C1 is the full historical series (1912 to 2019).
* C2 is the last ~30 years (from 1989), a drier, more pessimistic window.

```{r scenarios}
year <- as.integer(format(precip_pi$date, "%Y"))
p_c1 <- precip_pi$value
p_c2 <- precip_pi$value[year >= 1989]

sim_c1 <- rh_simulate(p_c1, demand = demand, area = area_total, capacity = 400,
                      runoff = 0.85, efficiency = 1, initial = 400)
sim_c2 <- rh_simulate(p_c2, demand = demand, area = area_total, capacity = 400,
                      runoff = 0.85, efficiency = 1, initial = 400)

sim_c1
```

## Annual results

Dividing the totals by the number of years gives the average annual performance
of the system under each scenario:

```{r annual}
n_years_c1 <- length(unique(year))
n_years_c2 <- length(unique(year[year >= 1989]))

results <- data.frame(
  Variable = c("Overflow (m3/yr)", "Deficit (m3/yr)",
               "Usable volume (m3/yr)", "Attendance (%)"),
  C1 = c(sim_c1$summary$overflow_total / n_years_c1,
         sim_c1$summary$deficit_total / n_years_c1,
         sim_c1$summary$usable_volume / n_years_c1,
         sim_c1$summary$attendance_pct),
  C2 = c(sim_c2$summary$overflow_total / n_years_c2,
         sim_c2$summary$deficit_total / n_years_c2,
         sim_c2$summary$usable_volume / n_years_c2,
         sim_c2$summary$attendance_pct)
)
print(results, digits = 5, row.names = FALSE)
```

With the current 400 m³ reservoir, about 70% of the demand is met in both
scenarios, and the large overflow points to the catchment as the bottleneck.

## Guarantee-based sizing

The *guaranteed demand* is the largest constant demand met with no deficit; the
*guaranteed capacity* is the smallest reservoir giving no deficit at the current
demand.

```{r sizing}
dgar_c1 <- rh_guaranteed_demand(p_c1, area = area_total, capacity = 400,
                                runoff = 0.85, efficiency = 1, initial = 400)
sgar_c1 <- rh_guaranteed_capacity(p_c1, demand = demand, area = area_total,
                                  runoff = 0.85, efficiency = 1)

c(guaranteed_demand_L_day = dgar_c1 * 1000,
  guaranteed_capacity_L   = sgar_c1 * 1000)
```

The guaranteed demand is about 1900 L/day (~29% of the campus demand), while
guaranteeing the full demand with the current catchment would require a reservoir
of several million litres. Both point to expanding the catchment area as the
main lever for water security on this campus.

## Reservoir behaviour

```{r autoplot, eval = has_ggplot}
# first three years, to see the reservoir dynamics
sim_zoom <- rh_simulate(p_c1[1:(3 * 365)], demand = demand, area = area_total,
                        capacity = 400, runoff = 0.85, efficiency = 1, initial = 400)
ggplot2::autoplot(sim_zoom)
```

## Per-block usable volume

The usable volume scales with catchment area, so which roofs are connected
matters for planning:

```{r per-block}
per_block <- data.frame(
  block = areas_pi$block,
  usable_m3_yr = vapply(areas_pi$area_m2, function(a) {
    s <- rh_simulate(p_c1, demand = demand, area = a, capacity = 400,
                     runoff = 0.85, efficiency = 1, initial = 400)
    s$summary$usable_volume / n_years_c1
  }, numeric(1))
)
per_block
```

## Reference

The IFPB-PI campus was studied in: Silva, R. M. V. da; Lourenço, A. M. G.;
Del Grande, M. H.; Farias, C. A. S. de; Albuquerque, E. M. de; Araújo, A. O. de
(2024). *Avaliação do potencial de aproveitamento de água de chuva usando
técnicas de modelagem hidrológica: estudo de caso do campus IFPB-PI.* XVII
Simpósio de Recursos Hídricos do Nordeste, João Pessoa-PB. ABRHidro.
